---
title: "NSF Grant Terminations"
description: Analysis of NSF grants terminated by the Trump Administration from TidyTuesday
author: Ava Li
datE: September 27, 2026
format: html
---

```{r}
#| echo: false
library(tidyverse)
library(openintro)
```
```{r}
nsf_terminations <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-05-06/nsf_terminations.csv')

nsf <- nsf_terminations
```

## The Data
Retrieved from TidyTuesday:
https://github.com/rfordatascience/tidytuesday/blob/main/data/2025/2025-05-06/readme.md

The data was collected by Grant Watch through crowdsourcing from researchers and program administrators, as the administration has not released information on these terminations.

The dataset includes information on grants for scientific research and education projects from the U.S. National Science Foundation (NSF) that have been terminated by the Trump administration in 2025.

```{r}
nsf |> 
    group_by(termination_letter_date) |> 
    summarize(num_perday = n()) |> 
    ggplot(aes(x=termination_letter_date,
        y=num_perday)) +
        geom_line() +
        geom_point() +
        labs(
            title = "NSF Grant Terminations Per Day",
            x = "Termination date",
            y = "Number of terminated grants")
```
```{r}
nsf |> 
    group_by(termination_letter_date) |> 
    summarize(num_perday = n()) |> 
    ggplot(aes(x=termination_letter_date,
        y=num_perday)) +
        geom_col() +
        labs(
            title = "NSF Grant Terminations Per Day",
            x = "Termination date",
            y = "Number of terminated grants")
```

```{r}
nsf |> 
    filter(!is.na(division)) |> 
    group_by(division) |>
    summarize(num = n()) |> 
    filter(num > 1) |> 
    ggplot(aes(x=num,y = fct_reorder(division, num)))+
        geom_col(fill = "#c84545") +
        labs(
            title = "NSF Grant Terminations by Division",
            x = "Number of Grants Terminated",
            y = NULL
        ) +
        theme_minimal()
```

```{r}
nsf |> 
    filter(!is.na(directorate)) |> 
    group_by(directorate) |>
    summarize(num = n()) |> 
    ggplot(aes(x=num,y = fct_reorder(directorate, num)))+
        geom_col(fill = "#c84545") +
        labs(
            title = "NSF Grant Terminations by Directorate",
            subtitle = "the NSF organization that administered the grant ",
            x = "Number of Grants Terminated",
            y = NULL
        ) +
        theme_minimal()
```

```{r}
nsf |> 
    group_by(org_state) |> 
    summarize(num = n()) |> 
    arrange(desc(num)) |> 
    slice_head(n = 16) |> 
    ggplot(aes(x=num,y = fct_reorder(org_state, num)))+
        geom_col(fill = "#2c7fb8") +
        labs(
            title = "NSF Grant Terminations by State",
            subtitle = "Top 16 States",
            x = "Number of Grants Terminated",
            y = NULL
        ) +
        theme_minimal()
```

```{r}
nsf |> 
    group_by(org_name) |> 
    summarize(num = n()) |> 
    filter(num>9) |> 
    ggplot(aes(x=num,y = fct_reorder(org_name, num)))+
        geom_col(fill = "#2c7fb8") +
        labs(
            title = "NSF Grant Terminations by Organization",
            subtitle = "10 or more terminations",
            x = "Number of Grants Terminated",
            y = NULL
        ) +
        theme_minimal()
```

```{r}
top_orgs <- nsf |>
  filter(!is.na(org_name), !is.na(usaspending_obligated)) |>
  group_by(org_name) |>
  summarize(total_dollars = sum(usaspending_obligated)) |>
  arrange(desc(total_dollars)) |>
  slice_head(n = 15)

ggplot(top_orgs, aes(x = total_dollars, y = fct_reorder(org_name, total_dollars))) +
  geom_col(fill = "#d95f0e") +
  scale_x_continuous(labels = scales::label_dollar(scale = 1e-6, suffix = "M")) +
  labs(
    title = "Organizations with the most NSF funding terminated",
    x = "Total obligated dollars",
    y = NULL
  ) +
  theme_minimal()
```

```{r}
org_summary <- nsf |>
  filter(!is.na(org_name), !is.na(usaspending_obligated)) |>
  group_by(org_name) |>
  summarize(
    n_terminations = n(),
    total_dollars = sum(usaspending_obligated)
  )

ggplot(org_summary, aes(x = n_terminations, y = total_dollars)) +
  geom_point(alpha = 0.6, color = "#2c7fb8") +
  scale_x_log10() +
  scale_y_log10(labels = scales::label_dollar(scale = 1e-6, suffix = "M")) +
  labs(
    title = "Terminated grants vs. dollars, by organization",
    x = "Number of terminated grants (log scale)",
    y = "Total obligated dollars (log scale)"
  ) +
  theme_minimal()
```
```{r}
nsf |> 
    group_by(org_district) |> 
    summarize(num = n()) |> 
    arrange(desc(num)) |> 
    slice_head(n = 20) |> 
    ggplot(aes(x=num,y = fct_reorder(org_district, num)))+
        geom_col(fill = "#b8622c") +
        labs(
            title = "NSF Grant Terminations by District",
            subtitle = "Top 20 Districts",
            x = "Number of Grants Terminated",
            y = NULL
        ) +
        theme_minimal()
```

```{r}
nsf |> 
    group_by(award_type) |> 
    summarize(num_term = n()) |> 
    mutate(perc = num_term/sum(num_term)) |> 
    ggplot(aes(x="",y = num_term, fill = award_type)) +
        geom_col() +
        coord_polar(theta = "y") +
        labs(
            title = "NSF Terminations by Award Type",
            fill = "Award Type"
        ) +
    theme_void()
```

```{r}
ggplot(nsf, aes(x=termination_letter_date, y=usaspending_obligated)) +
    geom_point()
```

