---
title: "NSF Grant Terminations"
description: Analysis of NSF grants terminated by the Trump Administration from TidyTuesday
author: Ava Li
datE: September 27, 2026
format: html
---

```{r}
#| echo: false
library(tidyverse)
library(openintro)
```

```{r}
nsf_terminations <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-05-06/nsf_terminations.csv')

nsf <- nsf_terminations
```

## The Data
Retrieved from TidyTuesday:
https://github.com/rfordatascience/tidytuesday/blob/main/data/2025/2025-05-06/readme.md

The data was collected by Grant Watch through crowdsourcing from researchers and program administrators, as the administration has not released information on these terminations.

The dataset includes information on grants for scientific research and education projects from the U.S. National Science Foundation (NSF) that have been terminated by the Trump administration in 2025.


```{r}
nsf <- nsf |> 
    group_by(termination_letter_date) |> 
    summarize(num = n())
```

Time Trends
```{r}
ggplot(nsf, aes(x=termination_letter_date)) +
    geom_line()

```